Capitalized Title Here

### 2021-02-11 CMPSpaper_writing/CMPSpaper_writing.Rmd
pictures
couple of citations

@@

The datasets we worked with come from the James Hamby Consecutively Rifled Ruger Barrel Study [cite?], and in particular, they are Hamby set 252 and Hamby set 44. For each Hamby set in the study, 35 bullets are fired by ten Ruger P-85 pistol barrels. 20 of the 35 bullets are known test bullets and 15 of them are unknown bullets. Each bullet has six land engraved areas (LEA) or land impressions, whose 3D topographical images were obtained and stored in the x3p format (XML 3-D Surface Profile). The x3p format provides a standard way of exchanging 2D and 3D profile data. It conforms to the ISO5436-2 standard [http://sourceforge.net/p/open-gps/mwiki/X3p/] and was adopted by the OpenFMC (Open Forensic Metrology Consortium), a group of firearm forensics researchers who contributes to the establishment of best practices of using metrology in forensic science. Hamby set 252 were scanned using a NanoFocus lens at 20x magnification with the scan resolution being 1.5625 \textmu m $\times$ 1.5625 \textmu m per pixel. Hamby set 44 were scanned ...[at CSAFE?]... Both Hamby set 252 and Hamby set 44 are publicly available on the NIST Ballistics Database Project.

The automated framework  used in this paper was proposed by Hare and Hofmann in 2017 [AUTOMATIC MATCHING OF BULLET LAND IMPRESSIONS?]. In order to obtain bullet signatures from the x3p files of bullet land impressions, we used `R` packages `x3ptools`[cite?] and `bulletxtrctr`[cite?]. `x3ptools` can be used to process x3p files. In the case of Hamby set 252, we need to convert the measurements of x3p files from meters to microns (`x3ptools::x3p_m_to_mum`) and rotate x3p files to make the long axis horizontal (`x3ptools::x3p_rotate`). Figure 1(a) shows a bullet land impression scan of a processed x3p file. The raised portion of the surface are groove impressions, and the vertical striation marks between the groove impressions are what we are interested in, called a land impression. The first step of obtaining the bullet signature is to extract a cross-section at a fixed height along the land impression. `bulletxtrctr::x3p_crosscut_optimize` can be used to determine an ideal height, and `bulletxtrctr::x3p_crosscut` can be used to acquire the crosscut data (`ccdata`). Figure 1(c) shows the typical pattern of the crosscut data, where "wedges" can be found on the left and right side and indicate the boundary between the land impression and the groove impression. `bulletxtrctr::cc_locate_grooves` is used to locate and chop these wedge areas so that only the curve of the land impression in the middle is kept. With the groove impressions excluded, we then use `bulletxtrctr::cc_get_signature` to smooth out the land impression curvature using LOESS (locally estimated scatter plot smoothing) [Cleveland 1979?] and obtain the bullet signature. Figure 1 illustrates this framework. Sometimes the algorithm in our framework cannot determine the ideal location of the crosscut or the grooves due to the deformation of a bullet, so we also developed a `R` Shiny App called `bulletinspectR` to identify and correct those errors. Note that NIST researchers used a different way of getting the bullet signatures in their paper [cite?].

@@

Once we obtain the bullet signatures, we can compare them using an algorithm that computes a similarity score. The congruent matching profile segments (CMPS) algorithm, developed by NIST for "objective comparison of striated tool marks" [original paper] in 2019, is one such algorithm. The main idea of the CMPS algorithm is that, we cut the reference profile (one of the two bullet signatures) into consecutive and non-overlapping basis segments and ask each basis segment to find the "best" registration position on the comparison profile (the other bullet signature) in terms of correlation values. Then the congruent registration position is the position where the most basis segments find it the "best", and the number of basis segments taking the congruent registration position is the CMPS score. The CMPS algorithm can assist firearm examiners with drawing a conclusion, but NIST's implementation of the CMPS algorithm was not released for the public. Thus, in this paper we present an open-source implementation of the CMPS algorithm so that everyone can use and improve it. With our implementation, one can compute the CMPS score of a comparison using the following code:

>>>>

## Background

@@

The datasets we worked with come from the James Hamby Consecutively Rifled Ruger Barrel Study [@brundage; @hamby; @Hamby:2019], and in particular, they are Hamby set 252 and Hamby set 44. For each Hamby set in the study, 35 bullets are fired by ten Ruger P-85 pistol barrels. 20 of the 35 bullets are known test bullets and 15 of them are unknown bullets. Each bullet has six land engraved areas (LEA) or land impressions, whose 3D topographical images were obtained and stored in the x3p format (XML 3-D Surface Profile). The x3p format provides a standard way of exchanging 2D and 3D profile data. It conforms to the ISO5436-2 standard [http://sourceforge.net/p/open-gps/mwiki/X3p/] and was adopted by the OpenFMC (Open Forensic Metrology Consortium), a group of firearm forensics researchers who contributes to the establishment of best practices of using metrology in forensic science. Hamby set 252 were scanned using a NanoFocus lens at 20x magnification with the scan resolution being 1.5625 \textmu m $\times$ 1.5625 \textmu m per pixel. Hamby set 44 were scanned ...[at CSAFE?]... Both Hamby set 252 and Hamby set 44 are publicly available on the NIST Ballistics Database Project.

The automated framework  used in this paper was proposed by @aoas. In order to obtain bullet signatures from the x3p files of bullet land impressions, we used `R` packages `x3ptools`[cite?] and `bulletxtrctr`[cite?]. `x3ptools` can be used to process x3p files. In the case of Hamby set 252, we need to convert the measurements of x3p files from meters to microns (`x3ptools::x3p_m_to_mum`) and rotate x3p files to make the long axis horizontal (`x3ptools::x3p_rotate`). Figure 1(a) shows a bullet land impression scan of a processed x3p file. The raised portion of the surface are groove impressions, and the vertical striation marks between the groove impressions are what we are interested in, called a land impression. The first step of obtaining the bullet signature is to extract a cross-section at a fixed height along the land impression. `bulletxtrctr::x3p_crosscut_optimize` can be used to determine an ideal height, and `bulletxtrctr::x3p_crosscut` can be used to acquire the crosscut data (`ccdata`). Figure 1(c) shows the typical pattern of the crosscut data, where "wedges" can be found on the left and right side and indicate the boundary between the land impression and the groove impression. `bulletxtrctr::cc_locate_grooves` is used to locate and chop these wedge areas so that only the curve of the land impression in the middle is kept. With the groove impressions excluded, we then use `bulletxtrctr::cc_get_signature` to smooth out the land impression curvature using LOESS (locally estimated scatter plot smoothing) [Cleveland 1979?] and obtain the bullet signature. Figure 1 illustrates this framework. Sometimes the algorithm in our framework cannot determine the ideal location of the crosscut or the grooves due to the deformation of a bullet, so we also developed a `R` Shiny App called `bulletinspectR` to identify and correct those errors. Note that NIST researchers used a different way of getting the bullet signatures in their paper [cite?].

@@

Once we obtain the bullet signatures, we can compare them using an algorithm that computes a similarity score. The congruent matching profile segments (CMPS) algorithm, developed by NIST for "objective comparison of striated tool marks" [@cmps] in 2019, is one such algorithm. The main idea of the CMPS algorithm is that, we cut the reference profile (one of the two bullet signatures) into consecutive and non-overlapping basis segments and ask each basis segment to find the "best" registration position on the comparison profile (the other bullet signature) in terms of correlation values. Then the congruent registration position is the position where the most basis segments find it the "best", and the number of basis segments taking the congruent registration position is the CMPS score. The CMPS algorithm can assist firearm examiners with drawing a conclusion, but NIST's implementation of the CMPS algorithm was not released for the public. Thus, in this paper we present an open-source implementation of the CMPS algorithm so that everyone can use and improve it. With our implementation, one can compute the CMPS score of a comparison using the following code:
